---
title: "FlexRL-RL-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FlexRL-RL-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(FlexRL)
pak::pak("robachowyk/FlexRL")
library(FlexRL)
options(warn=-1)
```

# fastLink: memory error / multilink: memory error

`FlexRL` is a package proposing flexible probabilistic **R**ecord **L**inkage with diagnostic tools for downstream inference on linked data. The record linkage algorithm `StEM()` uses a **St**ochastic **E**xpectation **M**aximisation approach to combine 2 data sources and outputs the set of records referring to the same entities.

More details on the [record linkage methodology](https://doi.org/10.1093/jrsssc/qlaf016); on the [false discovery proportion estimation](https://doi.org/10.1002/sim.70292); and on the potential risks of downstream inference on linked data (Robach et al., in preparation).

Several other algorithms for record linkage are available in R. The `link_with_*()` wrappers of FlexRL run them through one common interface, so that the same diagnostics and false discovery proportion (FDP) estimators apply to all of them: `link_with_FlexRL()`, `link_with_BRL()`, `link_with_fastLink()`, `link_with_multilink()`, `link_with_reclin2()`, `link_with_diyar()`, `link_with_fedmatch()`. Each wrapper takes the two encoded sources and a named list of arguments forwarded to the underlying package, and returns a list with `idxA`, `idxB` (row indices of the linked records in dataA and dataB) and `LinkScore` (the linkage scores, `NULL` when the package does not return any).

This vignette runs these wrappers on the same pipeline for:
- a regional subset of the [NLTCS](https://www.icpsr.umich.edu/web/NACDA/studies/9681/versions/V5) data (1982 and 1994 waves)
- a regional subset of the [SHIW](https://github.com/robachowyk/FlexRL-experiments/tree/main/SHIWApplication/SHIWData) data (2016 and 2020 waves)
- the full SHIW data with and without regional blocking

## One pipeline for all the wrappers

```{r}
wrapper_args <- function(method, prep_data) {
  PIVs <- names(prep_data$PIVs_config)
  switch(method,
         FlexRL    = list(data = prep_data, StEM_iter = 10, StEM_burnin = 5, gibbs_iter = 10, gibbs_burnin = 5, n_post_samp = 100),
         BRL       = list(flds = PIVs, types = rep("bi", length(PIVs)), nIter = 1000),
         fastLink  = list(varnames = PIVs, tol.em = 1e-06, threshold.match = 0.5, return.all = TRUE, n.cores = 1),
         multilink = list(types = rep("bi", length(PIVs)), breaks = stats::setNames(as.list(rep(NA, length(PIVs))), PIVs), duplicates = c(0, 0), n_iter = 200, max_cc_size = 50),
         reclin2   = list(on = PIVs, formula = stats::as.formula(paste("~", paste(PIVs, collapse = "+"))), type = "mpost", add = TRUE, variable = "selected", score = "mpost", threshold = 0.5),
         diyar     = list(probabilistic = TRUE, return_weights = TRUE),
         fedmatch  = list(by = PIVs, match_type = "multivar", unique_key_1 = "unique_key_A", unique_key_2 = "unique_key_B", multivar_settings = fedmatch::build_multivar_settings(compare_type = rep("indicator", length(PIVs)), wgts = rep(1 / length(PIVs), length(PIVs))))
         )
}

run_wrapper <- function(method, prep_data, arguments) {
  PIVs <- names(prep_data$PIVs_config)
  link_with <- get(paste0("link_with_", method), envir = asNamespace("FlexRL"))
  if (method == "diyar") {
    prep_data$encodedA <- prep_data$encodedA[, c(PIVs, "local_id", "source")]
    prep_data$encodedB <- prep_data$encodedB[, c(PIVs, "local_id", "source")]
  }
  tryCatch(link_with(prep_data$encodedA, prep_data$encodedB, arguments),
           error = function(e) {
             if (grepl("cannot allocate|memory|bad_alloc|too large|long vectors", conditionMessage(e), ignore.case = TRUE)) {
               message(method, ": memory error (", conditionMessage(e), "), results reported as NA")
               return(NULL)
             }
             stop(e)
           })
}

evaluate_linkage <- function(fit, true_pairs, threshold = 0.5) {
  if (is.null(fit)) return(rep(NA_real_, 5))
  delta_result <- data.frame(idxA = fit$idxA, idxB = fit$idxB)
  if (!is.null(fit$LinkScore)) delta_result <- delta_result[fit$LinkScore > threshold, , drop = FALSE]
  linked_pairs   <- do.call(paste, c(delta_result, list(sep = "_")))
  true_positive  <- length(intersect(linked_pairs, true_pairs))
  false_positive <- length(setdiff(linked_pairs, true_pairs))
  false_negative <- length(setdiff(true_pairs, linked_pairs))
  c(TP = true_positive, FP = false_positive, FN = false_negative,
    sensitivity = true_positive / (true_positive + false_negative),
    FDP = false_positive / (true_positive + false_positive))
}

results_rows <- c("TP", 
                  "FP", 
                  "FN", 
                  "sensitivity", 
                  "FDP",
                  "hat FDP score", 
                  "hat FDP score*", 
                  "hat FDP synth*",
                  "max SMD", 
                  "min support IoU", 
                  "max MMD")

run_all_wrappers <- function(prep_data, PIVs_type, methods, FDP_estimation = TRUE,
                             maxIter4CV = 3, n_repeats = 5, ...) {
  op <- options()
  on.exit(options(op))
  PIVs <- names(prep_data$PIVs_config)
  true_pairs <- do.call(paste, c(prep_data$true_pairs, list(sep = "_")))
  df_results <- data.frame(matrix(NA, nrow = length(results_rows), ncol = 0))
  rownames(df_results) <- results_rows

  naive_fit <- naive_linkage(PIVs = PIVs, encodedA = prep_data$encodedA, encodedB = prep_data$encodedB)
  df_results[1:5, "Naive"] <- evaluate_linkage(list(idxA = naive_fit[[1]], idxB = naive_fit[[2]]), true_pairs)

  for (method in methods) {
    arguments <- wrapper_args(method, prep_data)
    fit <- run_wrapper(method, prep_data, arguments)
    df_results[1:5, method] <- evaluate_linkage(fit, true_pairs)
    if (is.null(fit) || length(fit$idxA) == 0) next

    encodedA <- prep_data$encodedA
    encodedB <- prep_data$encodedB
    if (method == "diyar") {
      encodedA <- encodedA[, c(PIVs, "local_id", "source")]
      encodedB <- encodedB[, c(PIVs, "local_id", "source")]
    }
    diag <- do.call(RL_diagnostics, c(list(fit, encodedA, encodedB, PIVs, PIVs_type, PIVs,
                                           true_pairs = prep_data$true_pairs, FDP_estimation = FDP_estimation,
                                           RL_method = method, maxIter4CV = maxIter4CV, n_repeats = n_repeats),
                                      arguments))
    if (FDP_estimation) {
      df_results[6, method] <- diag$FDP_measures$curve$RL_FDP_score[1]
      df_results[7, method] <- diag$FDP_measures$curve$augmRL_FDP_score[1]
      df_results[8, method] <- diag$FDP_measures$curve$augmRL_FDP_synth[1]
    }
    df_results[9,  method] <- max(diag$discrepancy_measures[1, grep("smd", names(diag$discrepancy_measures))])
    df_results[10, method] <- min(diag$discrepancy_measures[1, grep("iou", names(diag$discrepancy_measures))])
    df_results[11, method] <- max(diag$discrepancy_measures[1, grep("mmd", names(diag$discrepancy_measures))])
  }
  df_results
}

show_results <- function(df_results) {
  cat("Record Linkage\n");      print(round(df_results[1:5, ], 3))
  cat("\nFDP estimation\n");    print(round(df_results[6:8, ], 3))
  cat("\nLinked divergence\n"); print(round(df_results[9:11, ], 3))
}

methods <- c("FlexRL", "BRL", "fastLink", "multilink", "reclin2", "diyar", "fedmatch")
```

We will fill a table of results throughout the vignette, gathering for each record linkage method: linkage performance, data quality diagnostics: FDP estimation, measures of divergence of linked data from sources.

## Regional subset of the NLTCS data

We use 4 **P**artially **I**dentifying **V**ariable**s** to link the records: sex, birth month and birth year are `stable`, the state code may change over time but we cannot model these changes so we set this variable as `flexible`.

```{r}
df1982 <- read.csv("ex-reg-NLTCS-82.csv", row.names = 1)
df1994 <- read.csv("ex-reg-NLTCS-94.csv", row.names = 1)

PIVs_config <- list(
  SEX       = list(dynamics = "stable",   bound_mistakes = c(0.10, 0.10), fix_mistakes = c(NA, NA)),
  DOB_MONTH = list(dynamics = "stable",   bound_mistakes = c(0.10, 0.10), fix_mistakes = c(NA, NA)),
  DOB_YEAR  = list(dynamics = "stable",   bound_mistakes = c(0.10, 0.10), fix_mistakes = c(NA, NA)),
  STATE     = list(dynamics = "flexible", bound_mistakes = c(NA, NA),     fix_mistakes = c(NA, NA)))
PIVs <- names(PIVs_config)
PIVs_type <- list(SEX = FALSE, DOB_MONTH = TRUE, DOB_YEAR = TRUE, STATE = FALSE)
```

`prepare_data()` removes the records for which the linking variables are outside of the common support, checks that PIVs are not strongly associated with each other (which may degrade record linkage performance), maps PIV values to natural numbers and encodes missing values as 0. The StEM algorithm denotes the smallest source as `A` and the largest as `B`. The true pairs are known from the identifier ID (they are indexed on the rows of `encodedA` and `encodedB`, after the support restriction).

```{r}
prep_data <- prepare_data(df1982, df1994, "1982", "1994", PIVs_config, same_mistakes = TRUE, uniq_id = "ID")

c(n_A = nrow(prep_data$encodedA), n_B = nrow(prep_data$encodedB), n_links = nrow(prep_data$true_pairs))

prep_data$n_values
```

```{R}
df_results_NLTCS <- run_all_wrappers(prep_data, PIVs_type, methods)
```

```{r}
show_results(df_results_NLTCS)
```

## Regional subset of the SHIW data

We use 4 **P**artially **I**dentifying **V**ariable**s** to link the records: sex and birth year are `stable`, education level and marital status are `flexible`.

```{r}
df2016 <- read.csv("ex-reg-SHIW-16.csv", row.names = 1)
df2020 <- read.csv("ex-reg-SHIW-20.csv", row.names = 1)

PIVs_config <- list(
  ANASCI = list(dynamics = "stable", bound_mistakes = c(0.10, 0.10), fix_mistakes = c(NA, NA)),
  SESSO  = list(dynamics = "stable", bound_mistakes = c(0.10, 0.10), fix_mistakes = c(NA, NA)),
  STUDIO = list(dynamics = "stable", bound_mistakes = c(0.10, 0.10), fix_mistakes = c(NA, NA)),
  STACIV = list(dynamics = "stable", bound_mistakes = c(0.10, 0.10), fix_mistakes = c(NA, NA)))
PIVs <- names(PIVs_config)
PIVs_type <- list(ANASCI = TRUE, SESSO = FALSE, STUDIO = FALSE, STACIV = FALSE)

prep_data <- prepare_data(df2016, df2020, "2016", "2020", PIVs_config, same_mistakes = TRUE, uniq_id = "ID")

c(n_A = nrow(prep_data$encodedA), n_B = nrow(prep_data$encodedB), n_links = nrow(prep_data$true_pairs))

prep_data$n_values
```

```{r}
df_results_SHIW <- run_all_wrappers(prep_data, PIVs_type, methods)
```

```{r}
show_results(df_results_SHIW)
```

## Full SHIW data

```{r}
df2016 <- read.csv("SHIW2016.csv", row.names = 1)
df2020 <- read.csv("SHIW2020.csv", row.names = 1)
df2016 <- df2016[!df2016$ID %in% df2016$ID[duplicated(df2016$ID)], ]
df2020 <- df2020[!df2020$ID %in% df2020$ID[duplicated(df2020$ID)], ]

PIVs_config_full <- c(PIVs_config,
  list(IREG = list(dynamics = "stable", bound_mistakes = c(0.10, 0.10), fix_mistakes = c(NA, NA))))
PIVs_full <- names(PIVs_config_full)
PIVs_type_full <- c(PIVs_type, IREG = FALSE)

prep_data_full <- prepare_data(df2016, df2020, "2016", "2020", PIVs_config_full, same_mistakes = TRUE, uniq_id = "ID")
true_pairs_full <- do.call(paste, c(prep_data_full$true_pairs, list(sep = "_")))

df_results_full <- data.frame(matrix(NA, nrow = 6, ncol = 0))
rownames(df_results_full) <- c("TP", "FP", "FN", "sensitivity", "FDP", "seconds")
for (method in methods) {
  t0 <- Sys.time()
  fit <- run_wrapper(method, prep_data_full, wrapper_args(method, prep_data_full))
  df_results_full[1:5, method] <- evaluate_linkage(fit, true_pairs_full)
  df_results_full[6, method] <- if (is.null(fit)) NA else round(as.numeric(Sys.time() - t0, units = "secs"))
}
write.csv(df_results_full, "SHIW-full-results.csv")
```














With PIVs that evolve over time, when modelling is possible, it is natural to parametrise the probability that underlying true values of the registered data for a pair of linked records coincide using a survival function. It is possible to use an exponential model, whose baseline hazard is constant in time; a Weibull model, whose baseline hazard is monotone in time; a Gompertz model, whose baseline hazard grows or decays exponentially in time; a piecewise-constant model, whose baseline hazard takes one value on each interval defined by given cut points; a custom model, supplied as a survival function S(X, alpha, times) together with a number of parameters and their starting values. These models may assume proportional hazards, accelerated failure time, proportional odds, additive hazards, ... If the covariates are categorical, make sure to encode them into dummies and drop one category (an intercept is included automatically). Since only the agreement of the true values at the observed time gap is used, the likelihood is the same for every model, -sum(Hequal * log(S) + (1 - Hequal) * log(1 - S)), and the parameters are estimated with stats::nlminb() at each M-step. 

```{r}
n_values  <- c(5, 6, 7, 12)
p_mistake <- list(V1 = c(0.02, 0.02), V2 = c(0.02, 0.02),
                  V3 = c(0.02, 0.02), V4 = c(0.02, 0.02))
p_missing <- list(V1 = c(0.005, 0.005), V2 = c(0.005, 0.005),
                  V3 = c(0.005, 0.005), V4 = c(0.005, 0.005))
cond_hazard_params <- list(V1 = c(), V2 = c(), V3 = c(), V4 = log(c(0.7, 0.6, 0.5)))

gen_data <- simulate_data(
  PIVs_config=PIVs_config, n_values=n_values, n_records=c(450, 500), 
  n_links=400, p_mistake=p_mistake, p_missing=p_missing, 
  cond_hazard_params=cond_hazard_params, enforce_estimability=TRUE,
  model_dynamics=survival_model("exponential")
)
```

We pre-process the data for the record linkage task: remove the records for which the linking variables are outside of the common support, check that PIVs are not strongly associated with each other (which may degrade record linkage performance), map PIV values to natural numbers and encode missing values as 0. The StEM algorithm denotes the smallest source as `A` and the largest as `B`.

```{r}
prep_data <- prepare_data(
  data1=gen_data$data1, data2=gen_data$data2, label1="1", label2="2", 
  PIVs_config=PIVs_config, same_mistakes=TRUE, uniq_id="entity_id",
  restrict_support_intersection=TRUE
)
```

```{r}
head(prep_data$encodedA)
```

```{r}
head(prep_data$encodedB)
```

For this example we know the true linkage structure.

```{r}
prep_data$true_pairs
true_pairs <- do.call(paste, c(prep_data$true_pairs, list(sep="_")))
```

The number of unique values per **PIV** gives information on their discriminating power:

```{r}
prep_data$n_values
```

We gauge how hard the record linkage task is using a naive linkage approach: matching records on exact agreement of identifying information.

```{r}
naive_fit <- naive_linkage(PIVs=PIVs, encodedA=prep_data$encodedA, encodedB=prep_data$encodedB)

df_results = data.frame( matrix(NA, nrow=11, ncol=0) )
rownames(df_results) = c("TP              ",
                         "FP              ",
                         "FN              ",
                         "FDP             ",
                         "sensitivity     ",
                         "hat FDP score   ",
                         "hat FDP score*  ",
                         "hat FDP synth*  ",
                         "max SMD         ",
                         "min support IoU ",
                         "max MMD         ")

linked_pairs    = do.call(paste, c(naive_fit, list(sep="_")))
true_positive   = length( intersect(linked_pairs, true_pairs) ) 
false_positive  = length( setdiff(linked_pairs, true_pairs) ) 
false_negative  = length( setdiff(true_pairs, linked_pairs) )
fdp             = false_positive / (true_positive + false_positive) 
sensitivity     = true_positive / (true_positive + false_negative) 
df_results[1:5,"Naive     "] = c(true_positive,false_positive,false_negative,sensitivity,fdp)

round(df_results, 2)
```

We run record linkage with `FlexRL`. The `data` should contain encodedA, encodedB, n_values, PIVs_config, same_mistakes. `StEM_iter`, `StEM_burnin` and `gibbs_iter`, `gibbs_burnin` are total number of iterations (including burn-in) and iterations to be discarded as burn-in. If `music_on` the algorithm opens a short tune in the browser when it finishes. You can `save_info_iter` (save environment information at each iteration) to `new_directory`. You can set starting values of the algorithm for parameters gamma and phi with `gamma0`, `phiA0`, `phiB0`, and change the number of samples drawn a posteriori to provide a linkage estimate with `n_post_samp` (set to 1000 as default).

```{r}
fit <- StEM(data = prep_data, StEM_iter = 30, StEM_burnin = 15,
           gibbs_iter = 20, gibbs_burnin = 10, music_on = FALSE)
```

The missing values and potential mistakes in the registration, the size of the overlapping set of records between the files, the total number of entities in the population, the low discriminative power of PIVs, PIVs potential dynamics across time, PIVs distribution, dependencies among PIVs, are obstacles to the linkage.

The algorithm returns: `Delta`, `gamma`, `eta`, `alpha`, `phi`.

```{r}
delta_result = fit$Delta
delta_result = delta_result[delta_result$x>0.5, ]

linked_pairs    = do.call(paste, c(delta_result[,c("i","j")], list(sep="_")))
true_positive   = length( intersect(linked_pairs, true_pairs) ) 
false_positive  = length( setdiff(linked_pairs, true_pairs) ) 
false_negative  = length( setdiff(true_pairs, linked_pairs) )
fdp             = false_positive / (true_positive + false_positive) 
sensitivity     = true_positive / (true_positive + false_negative) 

df_results[1:5,"FlexRL    "] = c(true_positive,false_positive,false_negative,sensitivity,fdp)

round(df_results, 2)
```

Let us run record linkage diagnostics:

```{r}
diag <- RL_diagnostics(fit = fit, encodedA = prep_data$encodedA, encodedB = prep_data$encodedB,
                      compare_vars = PIVs, vars_type_cont = PIVs_type, PIVs = PIVs,
                      true_pairs = prep_data$true_pairs, FDP_estimation = TRUE, 
                      RL_method = "FlexRL", data = prep_data,
                      maxIter4CV = 1, n_repeats = 2)

df_results[6,"FlexRL    "] = diag$FDP_measures$curve$RL_FDP_score[1]
df_results[7,"FlexRL    "] = diag$FDP_measures$curve$augmRL_FDP_score[1]
df_results[8,"FlexRL    "] = diag$FDP_measures$curve$augmRL_FDP_synth[1]

smd_idx = grep("smd", names(diag$discrepancy_measures))
iou_idx = grep("iou", names(diag$discrepancy_measures))
mmd_idx = grep("mmd", names(diag$discrepancy_measures))

df_results[9, "FlexRL    "] = max(diag$discrepancy_measures[1,smd_idx])
df_results[10,"FlexRL    "] = min(diag$discrepancy_measures[1,iou_idx])
df_results[11,"FlexRL    "] = max(diag$discrepancy_measures[1,mmd_idx])
```

```{r}
diag # print(diag)

print(diag, threshold = 0.75)
```

These summaries provide information on the linked data. We can also look at: the convergence of the StEM algorithm parameters, the linkage scores distribution, the data distributions between linked subset and original data sources, the divergence metrics of the linked data from the data sources, the FDP estimation.

```{r}
plot(diag, "convergence")

plot(diag, "scores")
 
plot(diag, "distributions", threshold = 0.75)

plot(diag, "discrepancy")

plot(diag, "FDP")
```

Several other algorithms for **R**ecord **L**inkage have been developed so far in the literature. Each has its own qualities and flaws. `StEM()` usually outperforms in scenarios where the **PIVs** have low discriminative power with few expected registration errors; it is particularly efficient when there is enough information to model the dynamics of some **PIV** changing across time (such as postal code). In 2026, it is the only package taking account of dynamic PIVs. More can be found on this topic in the simulation setting of the [methodology paper](https://doi.org/10.1093/jrsssc/qlaf016), code is provided in [`FlexRL-experiments`](https://github.com/robachowyk/FlexRL-experiments). More can be found on the FDP estimation methods in [this paper](https://doi.org/10.1002/sim.70292).

With [link_with_*()] wrappers we can link records with "multilink", "fastLink", "BRL", "reclin2", "diyar", "fedmatch".

```{r}
brl_args <- list(flds = PIVs, types = rep("bi",length(PIVs)))
fit_brl <- link_with_BRL(prep_data$encodedA, prep_data$encodedB, brl_args)

delta_result = as.data.frame(fit_brl)
delta_result = delta_result[delta_result$LinkScore>0.5, ]
linked_pairs = do.call(paste, c(delta_result[,c("idxA","idxB")], list(sep="_")))
true_positive = length( intersect(linked_pairs, true_pairs) )
false_positive = length( setdiff(linked_pairs, true_pairs) )
false_negative = length( setdiff(true_pairs, linked_pairs) )
fdp = false_positive / (true_positive + false_positive)
sensitivity = true_positive / (true_positive + false_negative)

df_results[1:5,"BRL       "] = c(true_positive,false_positive,false_negative,sensitivity,fdp)

diag_brl <- do.call(RL_diagnostics, c(list(fit_brl, prep_data$encodedA, prep_data$encodedB,
                           PIVs, PIVs_type, true_pairs = prep_data$true_pairs,
                           FDP_estimation = TRUE, RL_method = "BRL", maxIter4CV = 1, n_repeats = 2),
                           brl_args))

df_results[6,"BRL       "] = diag_brl$FDP_measures$curve$RL_FDP_score[1]
df_results[7,"BRL       "] = diag_brl$FDP_measures$curve$augmRL_FDP_score[1]
df_results[8,"BRL       "] = diag_brl$FDP_measures$curve$augmRL_FDP_synth[1]

smd_idx = grep("smd", names(diag_brl$discrepancy_measures))
iou_idx = grep("iou", names(diag_brl$discrepancy_measures))
mmd_idx = grep("mmd", names(diag_brl$discrepancy_measures))

df_results[9, "BRL       "] = max(diag_brl$discrepancy_measures[1,smd_idx])
df_results[10,"BRL       "] = min(diag_brl$discrepancy_measures[1,iou_idx])
df_results[11,"BRL       "] = max(diag_brl$discrepancy_measures[1,mmd_idx])
```

```{r}
multilink_args <- list(types = rep("bi",length(PIVs)), breaks = list(V1 = NA, V2 = NA, V3 = NA, V4 = NA),
                       duplicates = c(0,0), n_iter = 500, max_cc_size = 50)
fit_multilink <- link_with_multilink( prep_data$encodedA, prep_data$encodedB, multilink_args)

delta_result = as.data.frame(fit_multilink[c("idxA", "idxB")])
linked_pairs = do.call(paste, c(delta_result[,c("idxA","idxB")], list(sep="_")))
true_positive = length( intersect(linked_pairs, true_pairs) )
false_positive = length( setdiff(linked_pairs, true_pairs) )
false_negative = length( setdiff(true_pairs, linked_pairs) )
fdp = false_positive / (true_positive + false_positive)
sensitivity = true_positive / (true_positive + false_negative)

df_results[1:5,"multilink "] = c(true_positive,false_positive,false_negative,sensitivity,fdp)

diag_multilink <- do.call(RL_diagnostics, c(list(fit_multilink, prep_data$encodedA, prep_data$encodedB,
                                PIVs, PIVs_type, true_pairs = prep_data$true_pairs,
                                FDP_estimation = TRUE, RL_method = "multilink", maxIter4CV = 1, n_repeats = 2),
                                multilink_args))

df_results[6,"multilink "] = diag_multilink$FDP_measures$curve$RL_FDP_score[1]
df_results[7,"multilink "] = diag_multilink$FDP_measures$curve$augmRL_FDP_score[1]
df_results[8,"multilink "] = diag_multilink$FDP_measures$curve$augmRL_FDP_synth[1]

smd_idx = grep("smd", names(diag_multilink$discrepancy_measures))
iou_idx = grep("iou", names(diag_multilink$discrepancy_measures))
mmd_idx = grep("mmd", names(diag_multilink$discrepancy_measures))

df_results[9, "multilink "] = max(diag_multilink$discrepancy_measures[1,smd_idx])
df_results[10,"multilink "] = min(diag_multilink$discrepancy_measures[1,iou_idx])
df_results[11,"multilink "] = max(diag_multilink$discrepancy_measures[1,mmd_idx])
```

```{r}
fastLink_args <- list(varnames = PIVs, tol.em = 1e-06, threshold.match = 0.5, return.all = TRUE, n.cores = 1 )
fit_fastLink <- link_with_fastLink( prep_data$encodedA, prep_data$encodedB, fastLink_args )

delta_result = as.data.frame(fit_fastLink)
delta_result = delta_result[delta_result$LinkScore>0.5, ]
linked_pairs = do.call(paste, c(delta_result[,c("idxA","idxB")], list(sep="_")))
true_positive = length( intersect(linked_pairs, true_pairs) )
false_positive = length( setdiff(linked_pairs, true_pairs) )
false_negative = length( setdiff(true_pairs, linked_pairs) )
fdp = false_positive / (true_positive + false_positive)
sensitivity = true_positive / (true_positive + false_negative)

df_results[1:5,"fastLink  "] = c(true_positive,false_positive,false_negative,sensitivity,fdp)

diag_fastLink <- do.call(RL_diagnostics, c(list(fit_multilink, prep_data$encodedA, prep_data$encodedB,
                                PIVs, PIVs_type, true_pairs = prep_data$true_pairs,
                                FDP_estimation = TRUE, RL_method = "fastLink", maxIter4CV = 1, n_repeats = 2),
                                fastLink_args))

df_results[6,"fastLink  "] = diag_fastLink$FDP_measures$curve$RL_FDP_score[1]
df_results[7,"fastLink  "] = diag_fastLink$FDP_measures$curve$augmRL_FDP_score[1]
df_results[8,"fastLink  "] = diag_fastLink$FDP_measures$curve$augmRL_FDP_synth[1]

smd_idx = grep("smd", names(diag_fastLink$discrepancy_measures))
iou_idx = grep("iou", names(diag_fastLink$discrepancy_measures))
mmd_idx = grep("mmd", names(diag_fastLink$discrepancy_measures))

df_results[9, "fastLink  "] = max(diag_fastLink$discrepancy_measures[1,smd_idx])
df_results[10,"fastLink  "] = min(diag_fastLink$discrepancy_measures[1,iou_idx])
df_results[11,"fastLink  "] = max(diag_fastLink$discrepancy_measures[1,mmd_idx])
```

```{r}
reclin2_args <- list(on = PIVs, formula = stats::as.formula(paste("~", paste(PIVs, collapse = "+"))),
                     type = "mpost", add = TRUE, variable = "selected", score = "mpost", threshold = 0.5)
fit_reclin2 <- link_with_reclin2( prep_data$encodedA, prep_data$encodedB, reclin2_args )

delta_result = as.data.frame(fit_reclin2)
delta_result = delta_result[delta_result$LinkScore>0.5, ]
linked_pairs = do.call(paste, c(delta_result[,c("idxA","idxB")], list(sep="_")))
true_positive = length( intersect(linked_pairs, true_pairs) )
false_positive = length( setdiff(linked_pairs, true_pairs) )
false_negative = length( setdiff(true_pairs, linked_pairs) )
fdp = false_positive / (true_positive + false_positive)
sensitivity = true_positive / (true_positive + false_negative)

df_results[1:5,"reclin2   "] = c(true_positive,false_positive,false_negative,sensitivity,fdp)

diag_reclin2 <- do.call(RL_diagnostics, c(list(fit_reclin2, prep_data$encodedA, prep_data$encodedB, 
                               PIVs, PIVs_type, true_pairs = prep_data$true_pairs, 
                               FDP_estimation = TRUE, RL_method = "reclin2", maxIter4CV = 1, n_repeats = 2),
                               reclin2_args))

df_results[6,"reclin2   "] = diag_reclin2$FDP_measures$curve$RL_FDP_score[1]
df_results[7,"reclin2   "] = diag_reclin2$FDP_measures$curve$augmRL_FDP_score[1]
df_results[8,"reclin2   "] = diag_reclin2$FDP_measures$curve$augmRL_FDP_synth[1]

smd_idx = grep("smd", names(diag_reclin2$discrepancy_measures))
iou_idx = grep("iou", names(diag_reclin2$discrepancy_measures))
mmd_idx = grep("mmd", names(diag_reclin2$discrepancy_measures))

df_results[9, "reclin2   "] = max(diag_reclin2$discrepancy_measures[1,smd_idx])
df_results[10,"reclin2   "] = min(diag_reclin2$discrepancy_measures[1,iou_idx])
df_results[11,"reclin2   "] = max(diag_reclin2$discrepancy_measures[1,mmd_idx])
```

```{r}
diyar_args <- list(probabilistic = TRUE, return_weights = TRUE)
fit_diyar <- link_with_diyar(prep_data$encodedA[, PIVs], prep_data$encodedB[, PIVs], diyar_args)

delta_result = as.data.frame(fit_diyar[c("idxA", "idxB")])
linked_pairs = do.call(paste, c(delta_result[,c("idxA","idxB")], list(sep="_")))
true_positive = length( intersect(linked_pairs, true_pairs) )
false_positive = length( setdiff(linked_pairs, true_pairs) )
false_negative = length( setdiff(true_pairs, linked_pairs) )
fdp = false_positive / (true_positive + false_positive)
sensitivity = true_positive / (true_positive + false_negative)

df_results[1:5,"diyar     "] = c(true_positive,false_positive,false_negative,sensitivity,fdp)

names4diag <- c(PIVs, "local_id", "source")
diag_diyar <- do.call(RL_diagnostics, c(list(fit_diyar, prep_data$encodedA[, names4diag], prep_data$encodedB[, names4diag], 
                             PIVs, PIVs_type, true_pairs = prep_data$true_pairs, 
                             FDP_estimation = TRUE, RL_method = "diyar", maxIter4CV = 1, n_repeats = 2),
                             diyar_args))

df_results[6,"diyar     "] = diag_diyar$FDP_measures$curve$RL_FDP_score[1]
df_results[7,"diyar     "] = diag_diyar$FDP_measures$curve$augmRL_FDP_score[1]
df_results[8,"diyar     "] = diag_diyar$FDP_measures$curve$augmRL_FDP_synth[1]

smd_idx = grep("smd", names(diag_diyar$discrepancy_measures))
iou_idx = grep("iou", names(diag_diyar$discrepancy_measures))
mmd_idx = grep("mmd", names(diag_diyar$discrepancy_measures))

df_results[9, "diyar     "] = max(diag_diyar$discrepancy_measures[1,smd_idx])
df_results[10,"diyar     "] = min(diag_diyar$discrepancy_measures[1,iou_idx])
df_results[11,"diyar     "] = max(diag_diyar$discrepancy_measures[1,mmd_idx])
```

```{r}
fedmatch_args <- list(by = PIVs, match_type = "multivar", unique_key_1 = "unique_key_A", unique_key_2 = "unique_key_B",
                      multivar_settings = fedmatch::build_multivar_settings(compare_type = rep("indicator", length(PIVs)),
                                                                            wgts = rep(1/length(PIVs), length(PIVs)))
                      )
fit_fedmatch <- link_with_fedmatch( prep_data$encodedA, prep_data$encodedB, fedmatch_args )

delta_result = as.data.frame(fit_fedmatch)
delta_result = delta_result[delta_result$LinkScore>0.5, ]
linked_pairs = do.call(paste, c(delta_result[,c("idxA","idxB")], list(sep="_")))
true_positive = length( intersect(linked_pairs, true_pairs) )
false_positive = length( setdiff(linked_pairs, true_pairs) )
false_negative = length( setdiff(true_pairs, linked_pairs) )
fdp = false_positive / (true_positive + false_positive)
sensitivity = true_positive / (true_positive + false_negative)

df_results[1:5,"fedmatch  "] = c(true_positive,false_positive,false_negative,sensitivity,fdp)

diag_fedmatch <- do.call(RL_diagnostics, c(list(fit_fedmatch, prep_data$encodedA, prep_data$encodedB, 
                               PIVs, PIVs_type, true_pairs = prep_data$true_pairs, 
                               FDP_estimation = TRUE, RL_method = "fedmatch", maxIter4CV = 1, n_repeats = 2),
                               fedmatch_args))

df_results[6,"fedmatch  "] = diag_fedmatch$FDP_measures$curve$RL_FDP_score[1]
df_results[7,"fedmatch  "] = diag_fedmatch$FDP_measures$curve$augmRL_FDP_score[1]
df_results[8,"fedmatch  "] = diag_fedmatch$FDP_measures$curve$augmRL_FDP_synth[1]

smd_idx = grep("smd", names(diag_fedmatch$discrepancy_measures))
iou_idx = grep("iou", names(diag_fedmatch$discrepancy_measures))
mmd_idx = grep("mmd", names(diag_fedmatch$discrepancy_measures))

df_results[9, "fedmatch  "] = max(diag_fedmatch$discrepancy_measures[1,smd_idx])
df_results[10,"fedmatch  "] = min(diag_fedmatch$discrepancy_measures[1,iou_idx])
df_results[11,"fedmatch  "] = max(diag_fedmatch$discrepancy_measures[1,mmd_idx])
```

```{r}
options(digits = 2)

cat("Record Linkage ")
df_results[1:5,]

cat("\nFDP estimation ")
df_results[6:8,]

cat("\nLinked divergence ")
df_results[9:11,]
```
